\bf The Right Arm Bias in ITF Taekwon-Do \\ Patterns Chon-Ji to Juche

### 2012-02-25 article.tex
Added acknowledgements
Fixed spelling errors
Fixed spelling and grammar
Thanks Frøydis for your excellent feedback!

  When performing patterns in ITF Taekwon-Do, one of the factors mentioned in
  the encyclopedia is that there should be an equal amount of techniques to
  both sides of the body. Here we analyze hand techniques in Chon-Ji through
  Juche to establish whether or not there is a bias towards either side.  This
  is done by attaching metadata, in form of tags, to each pattern movement.
  These tags are extracted from each movement's English description in the
  condensed encyclopedia, and supplemented with keywords noted while
  performing the pattern. We then apply concepts from set theory to retrieve
  information we are interested in, and find that there are seven more
  techniques on the right side than on the left.

@@

  Combined in a pattern, it allows the student to systematically deal with one
  or several imaginary opponents under various assumptions, using every
  available attacking and blocking tool from different directions. In
  addition, the techniques should also be evenly distributed between the left
  and the right side\cite{cyclo:vol1}.

  It has, however, recently been observed that a bias towards the right leg
  exist. The study by Gibbs\cite{rlb} show that there is a total of 152 kicks
  in patterns Chon-Ji to Tong-Il. 85 kicks are with the right leg, and 67
  kicks are with the left leg. This means that 56 \% of the kicks are with the
  right leg, and 44 \% with the left leg.

  The work by Gibbs focused on kicks in all patterns, whereas this article
  examines patterns Chon-Ji to Juche. Looking at kicks in said patterns using
  Gibbs' paper, we find that there are 72 kicks in total. 42 kicks are with
  the right leg, and 30 kicks are with the left leg. Roughly 58 \% of the
  kicks are with the right leg, and about 42 \% with the left. Thus, the right
  leg bias still holds.

@@

  In a hierarchical and exclusive classification system, each object belong in
  one unambiguous category, which in turn is within another more general one.

@@

  Folder names are in themselves informative, in that, like tags, they
  describe the information held within them\cite{jones:folders}. Furthermore,
  the way folder names relate to each other show a structural relationship
  between directories\cite{mathes:folk}.

@@

  We then defined the data models with DataMapper, an Object Relational
  Mapper\footnote{\url{http://www.datamapper.org}} (ORM) written in the Ruby
  programming language. The models are Ruby classes with properties and
  associations. Properties define field names and data types in the database,
  whereas the associations defines the relationships and cardinality between
  the models.

@@

  Classification was done by attaching metadata, in terms of tags, to each
  pattern movement.

  The first movement in Chon-Ji, for instance, is to ``[m]ove the left foot to
  B, forming a left walking stance toward B while executing a low block to B
  with the left forearm\cite{cyclo:vol1}.'' Removing the verbose parts of the
  description leaves us with:

@@

  In most cases, applying tags to movements have been straightforward.
  However, there are movements that need special mention. These movements
  fall into two categories.

  The first category consist of movements where both arms are part of the
  technique by either forming the technique, or by being two separate
  techniques in a single movement.

  When both arms are used in one technique, the movement has been tagged with
  $\{left\_technique, \: right\_technique\}$ to reflect that both arms are
  needed for the technique, or that the movement contains two separate hand
  techniques.

@@

  The implications of tagging movements with $\{left\_technique, \:
  right\_technique\}$ is that the total amount of hand techniques on either
  side are higher than what actually is the case. This flaw does not affect
  the bias we are investigating, though.

  The second category comprises ambiguous movements, and thus contain
  clarification on why certain tags were added.

@@

        right\_technique\}$. This increases the total amount of blocks on both

@@

        right\_technique\}$. This increases the total amount of punches on

@@

        {\bf Movement 6}, \emph{Annun so ap joomuk noollo makgi}, consist of

@@

        {\bf Movement 7}, \emph{Annun so ap joomuk noollo makgi}, consist of

@@

        {\bf Movement 24}, \emph{Annun so ap joomuk noollo makgi}, consist of

@@

        {\bf Movement 25}, \emph{Annun so ap joomuk noollo makgi}, consist of

@@

        This increases the total amount of blocks on both sides.

@@

        has an effect on the total amount of low front snap kicks.

@@

        right\_technique\}$.  This increases the total amount of blocks on

@@

        right\_technique\}$. This increases the total amount of blocks on

@@

  In terms of category distribution, there are more blocks on the left side
  than there are on the right side. There are, therefore, more attacks on the
  right side, than on the left side (Figure \ref{fig:techniques_all}, Table
  \ref{tab:table2}).

  Independent on side or total, all technique categories follow the same
  pattern having most blocks, then punches, strikes and finally thrusts.

@@

  Techniques by height follow the same trend when looking at all techniques
  combined or techniques to either side. In general, there are more blocks
  than any other technique. Most of the techniques are in the middle section,
  although there is a greater occurrence of strikes in the high section, and
  very few attacks in the low section (Figure \ref{fig:height_all}).

@@

  Because most techniques are performed in walking, sitting and L stances, the
  graphs in this section have been split into two groups. One group shows the
  majority of techniques (84 \%), while the other visualize techniques
  performed in the remaining stances (16 \%).

@@

  There are some additional stances which needs special mentioning. The close
  stance is the only stance used by blocks, punches, strikes and thrusts. The
  parallel stance is only used for punches, whereas the vertical stance is
  only used for strikes. Only blocks are performed in the diagonal stance and
  the bending ready stance A. Finally, there are no techniques on the left
  side performed in the vertical stance.

  The total amount of techniques for both sides are 37 blocks and 34 attacks.
  Of all techniques on the left side, 59.5 \% are blocks, 40.46 \% are
  attacks. This means that there are 19.05 \% more blocks than attacks
  performed on that side. The right side follows the same pattern as the left,
  with more blocks than attacks. 57.51 \% of techniqes are blocks and 42.6 \%
  are attacks, meaning that there are 14.91 \% more blocks than attacks on the
  right side.

@@

  Excluding normal motion, it is interesting to note the high presence of
  blocks in continuous motion, which alone amounts to 34 \% of blocks
  performed (Figure \ref{fig:motion_all_not_n}). Continuing with Figure
  \ref{fig:motion_all_not_n}, it shows that there is an abnormal number of
  punches performed in fast motion compared to other motion types.

@@

  When examining the number of different motion types to either side, we find
  that there are more motion types on the right side, than on the left. The
  right side has flying motion and consecutive motion, in addition to motion
  types found on the left side. Also, there are more flying, slow, sliding and
  consecutive motions on the right side. However, there are more fast,
  stamping, continuous and connecting motions on the left side. Comparing the
  number of motion types on both sides reveals that there are equal amounts of
  slow, jumping, dodging, double stepping and shifting motions, although with
  different groups of techniques (Figures \ref{fig:motion_left_not_n},
  \ref{fig:motion_right_not_n}.

@@

  With a few exceptions, all techniques by all patterns for both sides, follow
  the same trend, where there are most blocks, then there are more punches
  than strikes, and more strikes than thrusts (Figure \ref{fig:patterns_all}).
  This trend is also true when grouping the patterns by black belt patterns
  and color belt patterns (Figures \ref{fig:patterns_black_all},
  \ref{fig:patterns_color_all}).

@@

  There are, however, patterns which don't follow the exceptions described
  above (Figures \ref{fig:patterns_left}, \ref{fig:patterns_right}). There
  are, for instance, more punches than blocks in patterns Chon-Ji and Dan-Gun
  on the right side than there are on the left side. And, when we compare the
  number of punches to strikes, both Toi-Gae and Choong-Moo have an equal
  amount of punches and strikes on the left side. Lastly, there is Choong-Jang
  with more strikes than punches on the right side.

@@

  There are more hand techniques performed on the right side than on the left
  side. Thusly, there is a bias towards the right side.

@@

% Unnumbered section (note the '*')
%\section*{Acknowledgments}

>>>>

  When performing patterns in ITF Taekwon-Do, one of the factors mentioned in
  the condensed Encyclopedia is that there should be an equal number of
  techniques to both sides of the body. Here we analyze hand techniques in
  Chon-Ji through Juche to establish whether or not there is a bias towards
  either side.  This is done by attaching metadata, in the form of tags, to
  each pattern movement. These tags are extracted from each movement's English
  description in the Encyclopedia, and supplemented with keywords noted while
  performing the pattern. We then apply concepts from set theory to retrieve
  the information we are interested in, and find that there are seven more
  techniques on the right side than on the left.

@@

  Combined in a pattern, it allows the student to systematically deal with one
  or several imaginary opponents using every available attacking and blocking
  tool from different directions. In addition, the techniques should also be
  evenly distributed between the left and the right side\cite{cyclo:vol1}.

  It has, however, recently been observed that there is a bias towards the
  right leg. The study by Gibbs\cite{rlb} show that there is a total of 152
  kicks in patterns Chon-Ji to Tong-Il. 85 kicks are with the right leg, and
  67 kicks are with the left leg. This means that 56 \% of the kicks are with
  the right leg, and 44 \% with the left leg.

  The work by Gibbs focused on kicks in all patterns, whereas this article
  examines patterns Chon-Ji to Juche. Looking at kicks in these patterns using
  Gibbs' paper, we find that there are 72 kicks in total. 42 kicks are with
  the right leg, and 30 kicks are with the left leg. Roughly 58 \% of the
  kicks are with the right leg, and about 42 \% with the left. Thus, the right
  leg bias still holds.

@@

  In a hierarchical and exclusive classification system, each object belongs
  to one unambiguous category, which in turn belongs within another more
  general category.

@@

  Folder names are in themselves informative, in that, like tags, they
  describe the information held within them\cite{jones:folders}. Furthermore,
  the way folder names relate to each other shows a structural relationship
  between directories\cite{mathes:folk}.

@@

  We then defined the data models with DataMapper, an Object Relational
  Mapper\footnote{\url{http://www.datamapper.org}} (ORM) written in the Ruby
  programming language. The models are Ruby classes with properties and
  associations. Properties define field names and data types in the database,
  whereas the associations define the relationships and cardinality between
  the models.

@@

  Classification was done by attaching metadata, in the form of tags, to each
  pattern movement.

  The first movement in Chon-Ji, for instance, is to ``[m]ove the left foot to
  B, forming a left walking stance toward B while executing a low block to B
  with the left forearm\cite{cyclo:vol1}.'' Removing the redundant parts of
  the description leaves us with:

@@

  In most cases, applying tags to movements has been straightforward.
  However, there are movements that need special mention. These movements fall
  into two categories.

  The first category consists of movements where both arms are part of the
  technique by either forming the technique, or by being two separate
  techniques in a single movement.

  In cases where both arms are used in one technique, the movement has been
  tagged with $\{left\_technique, \: right\_technique\}$ to reflect that both
  arms are needed for the technique, or that the movement contains two
  separate hand techniques.

@@

  The implications of tagging movements with $\{left\_technique, \:
  right\_technique\}$ are that the total amount of hand techniques on either
  side are higher than what actually is the case. This flaw does not affect
  the bias we are investigating, though.

  The second category comprises ambiguous movements, and thus contains
  clarification on why certain tags were added.

@@

        right\_technique\}$. This increases the total number of blocks on both

@@

        right\_technique\}$. This increases the total number of punches on

@@

        {\bf Movement 6}, \emph{Annun so ap joomuk noollo makgi}, consists of

@@

        {\bf Movement 7}, \emph{Annun so ap joomuk noollo makgi}, consists of

@@

        {\bf Movement 24}, \emph{Annun so ap joomuk noollo makgi}, consists of

@@

        {\bf Movement 25}, \emph{Annun so ap joomuk noollo makgi}, consists of

@@

        This increases the total number of blocks on both sides.

@@

        has an effect on the total number of low front snap kicks.

@@

        right\_technique\}$.  This increases the total number of blocks on

@@

        right\_technique\}$. This increases the total number of blocks on both

@@

  In terms of category distribution, there are more blocks on the left side
  than on the right side. There are, therefore, more attacks on the right
  side, than on the left side (Figure \ref{fig:techniques_all}, Table
  \ref{tab:table2}).

@@

  Techniques by height follow the same trend when looking at all techniques
  combined or techniques to either side. Most of the techniques are in the
  middle section, although there is a greater occurrence of strikes in the
  high section, and very few attacks in the low section (Figure
  \ref{fig:height_all}).

@@

  Because most techniques are performed in walking, sitting and L stances, the
  graphs in this section have been split into two groups. One group shows the
  majority of techniques (84 \%), while the other visualizes techniques
  performed in the remaining stances (16 \%).

@@

  There are some additional stances which needs special mentioning. The closed
  stance is the only stance used for blocking, punching, striking and
  thrusting. The parallel stance is only used for punching, whereas the
  vertical stance is only used for striking. Only blocks are performed in the
  diagonal stance and the bending ready stance A. Finally, there are no
  techniques on the left side performed in the vertical stance.

  The total number of techniques for both sides is 37 blocks and 34 attacks.
  Of all techniques on the left side, 59.5 \% are blocks, 40.46 \% are
  attacks. This means that there are 19.05 \% more blocks than attacks
  performed on that side. The right side follows the same pattern as the left,
  with more blocks than attacks. 57.51 \% of techniques are blocks and 42.6 \%
  are attacks, meaning that there are 14.91 \% more blocks than attacks on the
  right side.

@@

  Excluding normal motion, it is interesting to note the high presence of
  blocks in continuous motion, which alone amounts to 34 \% of blocks
  performed (Figure \ref{fig:motion_all_not_n}). Continuing with Figure
  \ref{fig:motion_all_not_n}, it shows that there is a high occurence of
  punches performed in fast motion compared to other motion types.

@@

  When examining the number of different motion types to either side, we find
  that there are more motion types on the right side, than on the left. The
  right side has flying motion and consecutive motion, in addition to motion
  types found on the left side. Also, there are more flying, slow, sliding and
  consecutive motions on the right side. However, there are more fast,
  stamping, continuous and connecting motions on the left side. Comparing the
  number of motion types on both sides reveals that there are equal amounts of
  slow, jumping, dodging, double stepping and shifting motions, although with
  different groups of techniques (Figures \ref{fig:motion_left_not_n},
  \ref{fig:motion_right_not_n}).

@@

  With a few exceptions, all techniques by all patterns for both sides, follow
  the same trend. There are more blocks than punches, more punches than
  strikes, and more strikes than thrusts (Figure \ref{fig:patterns_all}).
  This trend is also true when grouping the patterns by black belt patterns
  and color belt patterns (Figures \ref{fig:patterns_black_all},
  \ref{fig:patterns_color_all}).

@@

  There are, however, patterns which do not follow the exceptions described
  above (Figures \ref{fig:patterns_left}, \ref{fig:patterns_right}). There
  are, for instance, more punches than blocks in patterns Chon-Ji and Dan-Gun
  on the right side than there are on the left side. And, when we compare the
  number of punches to strikes, both Toi-Gae and Choong-Moo have an equal
  amount of punches and strikes on the left side. Lastly, there is Choong-Jang
  with more strikes than punches on the right side.

@@

  There are more hand techniques performed on the right side than on the left
  side. Thus, there is a bias towards the right side.

@@

% Unnumbered section (note the '*')
\section*{Acknowledgments}

  Todo.
